---
title: "An Example of Plotly in Web Presentations"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.align='center', message = FALSE, warning=FALSE)
```

## Completion Date: Jun 19, 2017

### Goal

This project will demonstrate the use of plotly in a slideshow. It will feature the oil dataset from the `caret` package and principal components analysis (PCA) to plot a 3D plot.

By plotting the first three principal components on a 3D plot, I expect to be able to clearly distinguish each type of oil as similar types should group together as the first three principal components will capture a large portion of the variability in the data.

## Load packages and data
Necessary packages:
```{r}
library(data.table)
library(caret)
library(plotly)
library(dplyr)
```
Load the data:
```{r}
data(oil)
oil_x = fattyAcids
oil_y = oilType
```



## Run PCA on the data

```{r}
train_pca_dat = oil_x
# Create preProcess object using pca method
train_pca_r <- preProcess(train_pca_dat, method = "pca", thresh = 0.99)
# Create the new dataset with pca variables from the preProcess object
train_pca <- predict(train_pca_r, train_pca_dat)
# Add the data you want to predict from training set
train_pca$oilType<- oil_y
```


## Code for the Plot
```{r, results="hide"}
plot_ly(x = train_pca$PC1, y = train_pca$PC2, z = train_pca$PC3, 
        color = oilType, sizes = 100, type="scatter3d",mode = "markers") %>% 
    layout(title = '3D plot for the first three principal components', 
           scene = list(xaxis = list(title = 'PC1'), 
                        yaxis = list(title = 'PC2'), 
                        zaxis = list(title = 'PC3')), 
           annotations = list(x = 1.13, y = 1.05, text = 'Oil Type',
                              xref = 'paper',yref = 'paper', 
                              showarrow = FALSE))
```

## The Plot
```{r, echo=FALSE}
plot_ly(x = train_pca$PC1, y = train_pca$PC2, z = train_pca$PC3, color = oilType, sizes = 100, type="scatter3d",mode = "markers") %>% layout(title = '3D plot for the first three principal components', scene = list(xaxis = list(title = 'PC1'), yaxis = list(title = 'PC2'), zaxis = list(title = 'PC3')), annotations = list(x = 1.13, y = 1.05,text = 'Oil Type',xref = 'paper',yref = 'paper', showarrow = FALSE))
```
Feel free to rotate the plot. As expected, the first three principal components do an excellent job of distinguishing the oil types!

## The End